One-at-a-time Parameter Perturbation Ensemble of the Community Land Model, version 5.1

### 2024-09-13 ms/ms.tex
overleaf

We repeated the full set of parameter perturbations across the six forcing scenarios in Table \ref{tab:exps} (Figure \ref{fig:ranges}). This ensures that we can identify parameters that are important not just under present-day conditions, but also parameters that control the response to forcing changes (CO$_2$, nitrogen deposition and climate). For example, the parameters that control present-day NEP differ  from the parameters that control NEP in the high CO$_2$ scenario (Figure \ref{fig:nep}a,b). NEP increases by approximately 50\% with the increase of CO$_2$ from 367 to 867 ppm when using the CLM5.1 default parameters, but can actually decrease with certain parameter settings (Figure \ref{fig:nep}c). One key parameter is \textit{tpuse\_sf}, which is a scalar perturbation factor influencing the triose phosphate limitations on photosynthesis. Triose phosphate limitation is difficult to quantify and has not been observed to significantly hinder photosynthesis under present-day CO$_2$ concentrations, but may play a larger role in the future \cite{lombardozzi2018,kumarathunge2019}. In CLM5.1, this parameter is not particularly influential on NEP at 367 ppm (ranked 39, not shown), but is third most influential when we increased CO$_2$ to 867 ppm.

The future climate experiment is 4.3K warmer averaged over our study domain (land-only, Antarctica excluded, Supp Figure S5). This led to an increase in evapotranspiration (ET) of 8.8\% relative to the control simulation when using the default CLM5.1 parameters (Figure \ref{fig:nep}f). While the parameters that most strongly affect global average ET in both the control and future climates are generally the same (Figure \ref{fig:nep}d,e), a distinct set of parameters govern the change in ET due to warming (Figure \ref{fig:nep}f). While the control simulation is primarily influenced by hydrology and stomatal conductance parameters, the response of ET to warming is largely controlled by photosynthesis acclimation parameters (i.e. \textit{tpuse\_sf}, \textit{kcha}, \textit{vcmaxha}, \textit{vcmaxhd}, \textit{lmrhd}, and \textit{cpha}). In a coupled ESM, land-atmosphere interactions would modulate these parameter effects, which could dampen parameter effects. While we feel relatively confident identifying influential parameters using land-only simulations, projecting their exact impacts in a coupled framework will require coupled model sensitivity tests.

@@

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{../figs/lai_biome.pdf}
\caption{The eight most influential parameters on leaf area index (LAI) within the control ensemble, globally and within three biomes. The solid lines span the range between the two simulations for each parameter, with an open circle indicating the simulation where a parameter is set to its minimum value and a filled circle for its maximum. The lines are color-coded according to scientific domain (Fig X). The dashed lines indicate the LAI for the simulation with default CLM5.1 parameters. Parameter rankings vary by biome, with the global rankings seeming to reflect contributions from each of these three biomes.}
\label{fig:lai}
\end{figure}

In this paper, we focus primarily on global and biome-level parameter rankings. However it is also possible to inspect the geographic footprint of parameter perturbations by projecting the sparsegrid output to standard lat/lon coordinates (see Section \ref{sect:sg} for details). For example, perturbing the \textit{medlynslope} parameter has a large effect on global runoff, but primarily via its effects in vegetated areas (Figure \ref{fig:map}). 
Because the number of potential variables, parameters, and geographical ranges of interest to the wider CLM community is larger than we can document here,  we provide a tool that can be used to explore an extended diagnostics set which summarises the $>$2TB of output data via approximately 2000 plots (\url{https://webext.cgd.ucar.edu/I2000/PPEn11_OAAT}). The diagnostics website includes ranking plots (as in Figure \ref{fig:lai}) and maps of parameter effects (as in Figure \ref{fig:map}). These plots are repeated across a combination of model output variables and model parameters. Likewise figures are repeated for each of the various forcing scenarios.

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{../figs/QRUNOFFabs_x_medlynslope_CTL2010.png}
\caption{Map of the effect of perturbing medlynslope on runoff within the control ensemble. Increasing medlynslope tends to reduce runoff, but only in regions with sufficient vegetation activity. This is one example of many plots available online (\url{https://webext.cgd.ucar.edu/I2000/PPEn11_OAAT}) in a broader diagnostics set. }
\label{fig:map}
\end{figure}

@@

In this project, we identified and perturbed 211 CLM parameters to create a large one-at-a-time PPE. There were several barriers to perturbing the full set of CLM-BGC parameters. First, many parameters had not been officially identified as parameters. In such cases, we identified hard-coded values, established an appropriate parameter name, and extracted that parameter to the CLM parameter file for easier manipulation. Although in this experiment we perturbed a large number of parameters across a variety of CLM processes (Figure \ref{fig:params}), this still does not cover the full set of CLM parameters, as some processes were not included, such as crop phenology and management. In carrying out this process of parameter identification, we likewise unearthed many nuances in the epistemology of what `parameters' are in the context of Earth system models. For instance, is soil texture a forcing variable or a parameter? Defining the parameters within a given model structure can be somewhat subjective, such that it would be difficult to collate a comprehensive or definitive set of parameters for a model like CLM. For example, for several empirical regressions, we scaled the slope and intercept terms in tandem, because we lacked sufficient information to define two independent perturbation ranges.

The second challenge involved defining a perturbation range for each parameter. We solicited expert judgment to set a minimum and maximum reasonable value for each parameter. In some cases, literature values were explicitly used (e.g. for the slope parameter of the stomatal conductance model, `medlynslope', \citeA{lin2015}), but the most common range was $\pm20\%$. It is exceedingly difficult to set parameter ranges that sample comparable probability density, even in a univariate experiment, like this one. The number of parameters is large, with many lacking sufficient empirical backing for robust range evaluation. Defining appropriate ranges for the full suite of land model parameters is a major activity of the scientific subdomains that comprise land models \cite{kattge2020}. Even with parameters that have an empirical basis from field studies, they may behave differently at the coarse spatial resolution used by climate models (zqz), complicating the process of uncertainty quantification. As such, we cannot claim that the parameters are equivalently sampled, whereby parameter effects and parameter rankings could be subject to sampling asymmetries.

@@

Just as the most influential parameters varied by forcing scenario, parameter rankings varied significantly by biome (Figure \ref{fig:lai}). Different PFTs present varying trait strategies, and different climates impose unique challenges to vegetation activity, such as light availability, extreme temperatures, or water stress. As such, perturbations that serve to ease or exacerbate the most substantial environmental stress for the set of PFTs in a given biome will have greater effect. Understanding the regional signatures of parameter perturbations is important for model calibration. Previous efforts have experienced difficulty resolving regional biases due to a limited amount of parameter flexibility. For example, \cite{dagon2020} could not simultaneously optimize plant productivity in the Amazon and Sahel regions. In some cases, perturbing PFT-specific parameters independently could aid in providing regional levers for bias reduction. In our experiments, PFT-specific parameters were perturbed in unison, but their effects can be analyzed by PFT for many variables that are modeled at the PFT-level, such as photosynthesis and transpiration. To first order, we expect that PFT-specific parameter effects will be independent, but CLM does feature some implicit competition among natural vegetation PFTs, such as through a shared soil moisture reservoir, which could impose some cross PFT interaction.

@@

Another caveat is that our PPE utilizes land-only simulations, with prescribed atmospheric forcing. This is likely to be adequate for estimating many land surface fluxes and pools, in particular of carbon, but precludes the influence of land-atmosphere feedbacks. This could lead to biased estimates of water and energy fluxes, as atmospheric feedbacks can modify the state of the land surface itself \cite{lague2019}. Additional work is ongoing to extend this work in a coupled model framework to quantify how much, where, and for which processes atmospheric feedbacks are likely to modulate land parameter effects substantially. Our expectation is that relative parameter rankings are likely robust, but that the absolute value of parameter effects may be significantly biased relative to their effects in the coupled model.

The primary deliverable of this project is the PPE dataset that captures the effects of 211 parameters on a wide range of CLM variables across six forcing scenarios. Several spin-off projects have leveraged this dataset, including some that have already reached publication \cite{cheng2023,yan2023a,yan2023b}. These projects utilized our ensemble to filter for parameters that influence the relevant study domains and variables, and performed follow-on experiments using the parameter ranges collated in Section \ref{sect:exps}.

The infrastructure underpinning this experiment is another major deliverable. Our project has accelerated parameter exploration work within our collaborator network by providing:

@@

We have already begun a follow-on activity that perturbs a subset of important parameters with a more exhaustive sampling approach to work towards calibration of leaf area index. By investing in the infrastructure that we introduce here, all of our subsequent parameter perturbation experiments have required much less time and effort. We continue to extend our efforts in this domain towards model calibration and uncertainty quantification, and we anticipate repeating the foundational one-at-a-time experiment with subsequent model releases. Reducing both the human and computer time required for these experiments is allowing us to repeat them more frequently through the model development cycle.

@@

A relatively small ($\sim$300MB), post-processed dataset is freely available online. All of the figures on our diagnostics website can be made from this one file. The raw dataset will be available on the NSF-NCAR computing system, and can be made available online as well, but we are gauging interest before committing to host the full 2TB dataset.

@@

\acknowledgments
This material is based upon work supported by the NSF National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No. 1852977. Computing and data storage resources, including the Cheyenne supercomputer (doi:10.5065/D6RX99HX) were provided by the Climate Simulation Laboratory at NSF-NCAR’s Computational and Information Systems Laboratory (CISL). DK and DML acknowledge funding from the National Oceanic and Atmospheric Administration (NOAA) Modelling, Analysis, Predictions and Projections (MAPP) program (award number NA20OAR4310413) and funding from the National Center for Atmospheric Research, which is a major facility sponsored by the NSF under Cooperative Agreement No. 1852977. KD is supported by the U.S. Department of Energy, Office of Science, Office of Biological \& Environmental Research (BER), Regional and Global Model Analysis (RGMA) component of the Earth and Environmental System Modeling Program under Award Number DE-SC0022070 and National Science Foundation (NSF) IA 1947282. ALSS recognizes support from the National Science Foundation (NSF) (AGS-1553715, AGS-2330096). FL is supported by the National Key Research and Development Program of China (2022YFE0106500). CDK acknowledges support by the Director, Office of Science, Office of Biological and Environmental Research of the US Department of Energy under contract DE-AC02-05CH11231 through the Regional and Global Model Analysis Program (RUBISCO SFA). RF and BS acknowledge funding by the European Union’s Horizon 2020 (H2020) research and innovation program under Grant Agreement No. 101003536 (ESM2025 – Earth System Models for the Future) and 821003 (4C, Climate-Carbon Interactions in the Coming Century).

@@

extra text:
Maintaining, improving, and interpreting complex land models benefits from thoughtful investment in software to automate and routinize important components of the development process, e.g. \citeA{collier2018}.

>>>>

We repeated the full set of parameter perturbations across the six forcing scenarios in Table \ref{tab:exps} (Figure \ref{fig:ranges}). This ensures that we can identify parameters that are important not just under present-day conditions, but also parameters that control the response to forcing changes (CO$_2$, nitrogen deposition and climate). For example, the parameters that control present-day NEP differ  from the parameters that control NEP in the high CO$_2$ scenario (Figure \ref{fig:nep}a,b). NEP increases by approximately 50\% with the increase of CO$_2$ from 367 to 867 ppm when using the CLM5.1 default parameters, but can actually decrease with certain parameter settings (Figure \ref{fig:nep}c). One key parameter is \textit{tpuse\_sf}, which is a scalar perturbation factor influencing the triose phosphate limitations on photosynthesis. Triose phosphate limitation is difficult to quantify and has not been observed to significantly hinder photosynthesis under present-day CO$_2$ concentrations, but may play a larger role in the future \cite{lombardozzi2018,kumarathunge2019}. This parameter is not particularly influential on NEP at 367 ppm (ranked 39, not shown), but is third most influential when we increased CO$_2$ to 867 ppm.

The future climate experiment is 4.3K warmer averaged over our study domain (land-only, Antarctica excluded, Supp Figure S5). This contributed to an increase in evapotranspiration (ET) of 8.8\% relative to the control simulation when using the default CLM5.1 parameters (Figure \ref{fig:nep}f). While the parameters that most strongly affect global average ET in both the control and future climates are generally the same (Figure \ref{fig:nep}d,e), a distinct set of parameters govern the change in ET due to warming (Figure \ref{fig:nep}f). While the control simulation is primarily influenced by hydrology and stomatal conductance parameters, the response of ET to warming is largely controlled by photosynthesis acclimation parameters (i.e. \textit{tpuse\_sf}, \textit{kcha}, \textit{vcmaxha}, \textit{vcmaxhd}, \textit{lmrhd}, and \textit{cpha}). In a coupled ESM, land-atmosphere interactions would modulate these parameter effects, which could dampen parameter effects. While we feel relatively confident identifying influential parameters using land-only simulations, projecting their exact impacts in a coupled framework will require coupled model sensitivity tests.

@@

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{../figs/lai_biome.pdf}
\caption{The eight most influential parameters on leaf area index (LAI) within the control ensemble, globally and within three biomes. The solid lines span the range between the two simulations for each parameter, with an open circle indicating the simulation where a parameter is set to its minimum value and a filled circle for its maximum. The lines are color-coded according to scientific domain. The dashed lines indicate the LAI for the simulation with default CLM5.1 parameters. Parameter rankings vary by biome, with the global rankings seeming to reflect contributions from each of these three biomes.}
\label{fig:lai}
\end{figure}

In this paper, we focus primarily on global and biome-level parameter rankings. However it is also possible to inspect the geographic footprint of parameter perturbations by projecting the sparsegrid output to standard lat/lon coordinates (see Section \ref{sect:sg} for details). For example, perturbing the \textit{medlynslope} parameter (which controls stomatal conductance) has a large effect on global runoff, but primarily via its effects in vegetated areas (Figure \ref{fig:map}). 
Because the number of potential variables, parameters, and geographical ranges of interest to the wider CLM community is larger than we can document here,  we provide a tool that can be used to explore an extended diagnostics set which summarises the $>$2TB of output data via approximately 2000 plots (\url{https://webext.cgd.ucar.edu/I2000/PPEn11_OAAT}). The diagnostics website includes ranking plots (as in Figure \ref{fig:lai}) and maps of parameter effects (as in Figure \ref{fig:map}). These plots are repeated across a combination of model output variables and model parameters. Likewise figures are repeated for each of the various forcing scenarios.

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{../figs/QRUNOFFabs_x_medlynslope_CTL2010.png}
\caption{Map of the effect of perturbing the medlynslope stomatal conductance parameter on runoff within the control ensemble. Increasing medlynslope tends to reduce runoff, but only in regions with sufficient vegetation activity. This is one example of many plots available online (\url{https://webext.cgd.ucar.edu/I2000/PPEn11_OAAT}) in a broader diagnostics set. }
\label{fig:map}
\end{figure}

@@

In this project, we identified and perturbed 211 CLM parameters to create a large one-at-a-time PPE. There were several barriers to perturbing the full set of CLM-BGC parameters. First, many parameters had not been officially identified as parameters. In such cases, we identified hard-coded values, established an appropriate parameter name, and extracted that parameter to the CLM parameter file for easier manipulation. Though we perturbed a large number of parameters across a variety of CLM processes (Figure \ref{fig:params}), this still does not cover the full set of CLM parameters, as some processes were not included, such as crop phenology and management. In carrying out this process of parameter identification, we likewise unearthed many nuances in the epistemology of what `parameters' are in the context of Earth system models. For instance, is soil texture a forcing variable or a parameter? Defining the parameters within a given model structure can be somewhat subjective, such that it would be difficult to collate a comprehensive or definitive set of parameters for a model like CLM. For example, for several empirical regressions, we scaled the slope and intercept terms in tandem, because we lacked sufficient information to define two independent perturbation ranges.

The second challenge involved defining a perturbation range for each parameter. We solicited expert judgment to set a minimum and maximum reasonable value for each parameter. In some cases, literature values were explicitly used (e.g. for the slope parameter of the stomatal conductance model, `medlynslope', \citeA{lin2015}), but the most common range was $\pm20\%$. It is exceedingly difficult to set parameter ranges that sample comparable probability density, even in a univariate experiment, like this one. The number of parameters is large, with many lacking sufficient empirical backing for robust range evaluation. Defining appropriate ranges for the full suite of land model parameters is a major activity of the scientific subdomains that comprise land models \cite{kattge2020}. Even with parameters that have an empirical basis from field studies, uncertainty quantification is difficult to the scale mismatch between site-level observations and land model grid cells \cite{liu2021}. As such, we cannot claim that the parameters are equivalently sampled, whereby parameter effects and parameter rankings could be subject to sampling asymmetries.

@@

Just as the most influential parameters varied by forcing scenario, parameter rankings varied significantly by biome (Figure \ref{fig:lai}). Different PFTs present varying trait strategies, and different climates impose unique challenges to vegetation activity, such as light availability, extreme temperatures, or water stress. As such, perturbations that serve to ease or exacerbate the most substantial environmental stress for the set of PFTs in a given biome will have greater effect. For example, modifying photosynthetic capacity (via jmaxb0, which sets baseline proportion of nitrogen allocated for electron transport) has a large effect in boreal forests, as light is relatively limiting at high latitudes.  Understanding the regional signatures of parameter perturbations is important for model calibration. Previous efforts have experienced difficulty resolving regional biases due to a limited amount of parameter flexibility. For example, \cite{dagon2020} could not simultaneously optimize plant productivity in the Amazon and Sahel regions. In some cases, perturbing PFT-specific parameters independently could aid in providing regional levers for bias reduction. In our experiments, PFT-specific parameters were perturbed in unison, but their effects can be analyzed by PFT for many variables that are modeled at the PFT-level, such as photosynthesis and transpiration. To first order, we expect that PFT-specific parameter effects will be independent, but CLM does feature some implicit competition among natural vegetation PFTs, such as through a shared soil moisture reservoir, which could impose some cross PFT interaction.

@@

Another caveat is that our PPE utilizes land-only simulations, with prescribed atmospheric forcing. This is likely to be adequate for estimating many land surface fluxes and pools, in particular of carbon, but precludes the influence of land-atmosphere feedbacks. This could lead to biased estimates of water and energy fluxes, as atmospheric feedbacks can modify the state of the land surface itself \cite{lague2019}. Additional work is ongoing to extend this work in a coupled model framework to quantify how much, where, and for which processes atmospheric feedbacks are likely to modulate land parameter effects substantially. 
Furthermore our experiments are conducted at nominal 2-degree resolution, certain processes subject to land heterogeneity may respond differently at higher resolution, such as permafrost depth in the Arctic \cite{schickhoff2024}.
Our expectation is that relative parameter rankings are likely robust, but that the absolute value of parameter effects may be significantly biased relative to their effects in the coupled model.

The primary deliverable of this project is the PPE dataset that captures the effects of 211 parameters on a wide range of CLM variables across six forcing scenarios. Several spin-off projects have leveraged this dataset, including some that have already reached publication \cite{cheng2023,yan2023a,yan2023b}. These projects utilized our ensemble to filter for parameters that influence the relevant study domains and variables, and performed follow-on experiments using the parameter ranges collated in Section \ref{sect:exps}. 
Maintaining, improving, and interpreting complex land models benefits from thoughtful investment in software to automate and routinize important components of the development process, such as the International Land Model Benchmarking (ILAMB) system \cite{collier2018}.
Therefore we also see the infrastructure underpinning our experiment as a major deliverable. Our project has accelerated parameter exploration work within our collaborator network by providing:

@@

We have already started a follow-on activity that perturbs a subset of important parameters with a more exhaustive sampling approach to work towards calibration of leaf area index. By investing in the infrastructure that we introduce here, all of our subsequent parameter perturbation experiments have required much less time and effort. We continue to extend our efforts in this domain towards model calibration and uncertainty quantification, and we anticipate repeating this foundational one-at-a-time experiment with subsequent model releases. Reducing both the human and computer time required for these experiments is allowing us to repeat them more frequently through the model development cycle.

@@

A relatively small ($\sim$300MB), post-processed dataset is freely available online (n.b. we will create a DOI and link accordingly once our figures are finalized later in the review process). All of the figures on our diagnostics website (\url{https://webext.cgd.ucar.edu/I2000/PPEn11_OAAT}) can be made from this one file. The raw dataset will be available on the NSF-NCAR computing system, and can be made available online as well, but we are gauging interest before committing to host the full 2TB dataset.

@@

\acknowledgments
This material is based upon work supported by the NSF National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No. 1852977. Computing and data storage resources, including the Cheyenne supercomputer (doi:10.5065/D6RX99HX) were provided by the Climate Simulation Laboratory at NSF-NCAR’s Computational and Information Systems Laboratory (CISL). DK and DML acknowledge funding from the National Oceanic and Atmospheric Administration (NOAA) Modelling, Analysis, Predictions and Projections (MAPP) program (award number NA20OAR4310413) and funding from the NSF National Center for Atmospheric Research, which is a major facility sponsored by the NSF under Cooperative Agreement No. 1852977. KD is supported by the U.S. Department of Energy, Office of Science, Office of Biological \& Environmental Research (BER), Regional and Global Model Analysis (RGMA) component of the Earth and Environmental System Modeling Program under Award Number DE-SC0022070 and National Science Foundation (NSF) IA 1947282. ALSS recognizes support from the National Science Foundation (NSF) (AGS-1553715, AGS-2330096). FL is supported by the National Key Research and Development Program of China (2022YFE0106500). CDK acknowledges support by the Director, Office of Science, Office of Biological and Environmental Research of the US Department of Energy under contract DE-AC02-05CH11231 through the Regional and Global Model Analysis Program (RUBISCO SFA). RF and BS acknowledge funding by the European Union’s Horizon 2020 (H2020) research and innovation program under Grant Agreement No. 101003536 (ESM2025 – Earth System Models for the Future) and 821003 (4C, Climate-Carbon Interactions in the Coming Century).

@@
